---
title: "Lab 2: Descriptive Statistics, Cross-Tabulations, and Joins"
subtitle: "In-Person Sections"
format:
  typst:
    toc: false
    papersize: us-letter
    fontsize: 10pt
    linestretch: 1
    margin:
      x: 1in
      y: 1in
    number-sections: false
    page-numbering: "1"
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(eco230r)
library(knitr)
options(scipen = 999)
```

## Purpose

Use two related NFL tables to practice identifying grain, checking keys,
joining data, and reproducing descriptive summaries across Excel, Tableau, and
R. The extract contains 2025 Minnesota regular-season games and their plays.

## 1) Load and inspect the tables

```{r}
games <- read_csv("data/nfl_games.csv", show_col_types = FALSE)
plays <- read_csv("data/nfl_plays.csv", show_col_types = FALSE)

glimpse(games)
glimpse(plays)
```

State the grain of each table and identify its key.

*Your response:*

## 2) Verify the relationship

Confirm that `game_id` is unique in `games` and repeated in `plays`. Explain
what those results imply about the relationship between the tables.

```{r}
games |> count(game_id) |> count(n)
plays |> count(game_id) |> arrange(desc(n))
```

*Your response:*

## 3) Join game information to plays

Use `left_join()` to add selected game-level fields. Verify that the join does
not change the number of play rows and investigate unmatched keys.

```{r}
plays_joined <- plays |>
  left_join(
    games |> select(game_id, week, game_date, home_team, away_team, total_points),
    by = "game_id"
  )

tibble(
  before = nrow(plays),
  after = nrow(plays_joined),
  unmatched = sum(is.na(plays_joined$week))
)
```

## 4) Build descriptive summaries

Create at least one count, one percentage, and one appropriate quantitative
summary by a categorical variable. Reproduce one summary in Excel and Tableau.

```{r}
plays_joined |>
  filter(!is.na(play_type)) |>
  group_by(play_type) |>
  summarise(
    plays = n(),
    mean_yards = mean(yards_gained, na.rm = TRUE),
    success_rate = mean(successful_play, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(plays)) |>
  kable(digits = 2)
```

## 5) Reconcile the tools

Compare the R, Excel, and Tableau results. If they differ, check filters,
missing-value handling, aggregation, and the join before deciding which result
is correct.

*Your response:*

## Completion checklist

- [ ] I stated each table's grain and key.
- [ ] I verified the join without multiplying play rows.
- [ ] I created a count, percentage, and quantitative summary.
- [ ] I reproduced and reconciled a summary across the required tools.
- [ ] I rendered `Lab_02_IP.pdf`.
