---
title: "Lab 2: Descriptive Statistics, Cross-Tabulations, and Joins"
subtitle: "In-Person Sections"
format:
  typst:
    toc: false
    papersize: us-letter
    fontsize: 10pt
    linestretch: 1
    margin:
      x: 1in
      y: 0.9in
    number-sections: false
    page-numbering: "1"
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(eco230r)
library(knitr)
options(scipen = 999)
```

## Purpose

Use two related NFL tables to practice joining data and producing the same
descriptive analysis in Excel, Tableau, and R. The extract contains 17
Minnesota regular-season games from 2025 and 2,791 plays from those games.

By the end of the lab, you should be able to:

- identify the grain and key of each table;
- explain how `XLOOKUP` or a relational join adds game information to plays;
- create counts and a frequency table;
- calculate and compare means and medians;
- create a cross-tabulation with meaningful percentages; and
- explain how missing values and repeated game-level values affect a summary.

## Files

- `data/Lab_02_Excel_Starter.xlsx` contains `Plays` and `Games` worksheets.
  The game fields are already joined to the right side of `Plays` with
  `XLOOKUP` formulas so you can inspect how the join works.
- `data/nfl_plays.csv` and `data/nfl_games.csv` contain the same two tables as
  individual files for Tableau and R.

## 1) Understand the two tables and their join

- `plays` has one row per play. The combination of `game_id` and `play_id` is
  the unique key.
- `games` has one row per game. Its unique key is `game_id`.
- The tables have a many-to-one relationship: many plays belong to one game.
- In Excel, inspect an existing `XLOOKUP` formula after the original play
  columns. Identify its lookup value, lookup array, and return array.
- Verify several returned values against the `Games` worksheet.

::: {.callout-caution appearance="minimal" icon="false" title="Keep the grain in mind"}
Joining game fields to plays repeats the same game-level value on every play in
that game. Do not sum final scores, total points, or other game-level values in
the joined play table. That would count each game's value once per play.
:::

Load and verify the two tables in R:

```{r}
games <- read_csv("data/nfl_games.csv", show_col_types = FALSE)
plays <- read_csv("data/nfl_plays.csv", show_col_types = FALSE)

games |> summarise(rows = n(), unique_games = n_distinct(game_id))

plays |>
  summarise(
    rows = n(),
    unique_game_play_keys = n_distinct(game_id, play_id)
  )
```

Join all game fields to the plays table and confirm that the number of play
rows does not change:

```{r}
plays_joined <- plays |>
  left_join(games, by = "game_id")

tibble(
  before = nrow(plays),
  after = nrow(plays_joined),
  unmatched_games = sum(is.na(plays_joined$season))
)
```

## 2) Create counts and a frequency table

In Excel, convert each source range to a Table with **Insert > Table**, confirm
that the range has headers, and name the tables `Plays` and `Games`. Then insert
a PivotTable from `Plays` on a new worksheet and rename the worksheet `Counts`.
Keep one PivotTable per worksheet so a refresh or layout change does not
overwrite another analysis.

- Add `play_type` to Rows.
- Add `game_id` to Values and summarize it by **Count**.
- Add `game_id` to Values a second time. Use **Show Values As** to display a
  percentage of the column total. The count and percentage together form a
  frequency table.
- Compare **Count of penalty_yards**, **Count of penalty**, and a summary of
  the `penalty` indicator. Explain how blanks and `FALSE` values affect each
  result.
- Filter `play_type` to `no_play` and inspect the play descriptions. Determine
  what distinguishes the records with a penalty from those without one.

The same frequency table in R is:

```{r}
play_frequency <- plays_joined |>
  filter(!is.na(play_type)) |>
  count(play_type, name = "plays") |>
  mutate(percent = plays / sum(plays)) |>
  arrange(desc(plays))

play_frequency |> kable(digits = 3)
```

Use a populated key when counting all rows. Counting a field with missing
values answers a different question:

```{r}
plays_joined |>
  summarise(
    all_plays = n(),
    populated_penalty_yards = sum(!is.na(penalty_yards)),
    penalties = sum(penalty, na.rm = TRUE)
  ) |>
  kable()
```

## 3) Compare means and medians

Create another PivotTable on a new worksheet named `Means`.

- Add `play_type` to Rows and `quarter` to Columns.
- Add `yards_gained` to Values and change the calculation to **Average**.
- Format the results as numbers with two decimal places.
- Filter `play_type` to `pass` and `run`, where yards gained is meaningful.
- Add `shotgun` below `play_type` in Rows, then replace it with
  `pass_location`. Look for expected patterns and results that need more
  investigation.
- Briefly drag `yardline_100` or `play_id` to Rows or Columns. Explain why a
  high-cardinality numeric field produces an unwieldy table rather than a
  useful summary.

Excel PivotTables do not offer median as a standard Value Field Setting. To
calculate a median, insert a new PivotTable with **Add this data to the Data
Model** selected, then create a measure such as:

```text
Med Yards := MEDIAN(Plays[yards_gained])
```

The median is direct in R:

```{r}
yard_summaries <- plays_joined |>
  filter(play_type %in% c("pass", "run")) |>
  group_by(quarter, play_type) |>
  summarise(
    plays = n(),
    mean_yards = mean(yards_gained, na.rm = TRUE),
    median_yards = median(yards_gained, na.rm = TRUE),
    .groups = "drop"
  )

yard_summaries |> kable(digits = 2)
```

## 4) Create a cross-tabulation

Create a third PivotTable on a new worksheet named `Cross Tab`.

- Add `play_type` to Rows and `quarter` to Columns.
- Add `game_id` to Values and summarize it by **Count**.
- Use **Show Values As** to compare percent of grand total, percent of column
  total, and percent of row total.
- For each version, state the denominator in words before interpreting a
  percentage.

Use percentages primarily with counts. Applying **Show Values As** to a mean
can create percentages that do not add to 100% and are difficult to explain.

Use `csi()` from `eco230r` to create the complete cross-tabulation. Its returned
object contains the observed counts plus row, column, and table percentages:

```{r}
library(eco230r)

csi_output <- plays_joined |>
  csi(play_type ~ quarter)

csi_output$observed |> kable(digits = 0)

csi_output$row_percentages |> kable(digits = 2)

csi_output$column_percentages |> kable(digits = 2)

csi_output$table_percentages |> kable(digits = 2)
```

## 5) Reproduce the analysis in Tableau

If Tableau is available:

- Connect to both CSV files and relate them on `game_id`.
- Verify that the relationship preserves the play-level grain.
- Create a text table with `play_type` on Rows and `COUNT(game_id)` on Text.
- Apply a Percent of Total table calculation and verify its computation
  direction before interpreting the denominator.
- Create a second table with `play_type` on Rows, `quarter` on Columns, and
  `AVG(yards_gained)` or `MEDIAN(yards_gained)` on Text.
- Reproduce the play-type-by-quarter cross-tabulation and compare it with the
  Excel and R results.

## 6) Reconcile the tools

Compare the Excel, Tableau, and R results. If they differ, check:

- filters and included categories;
- whether missing values were counted or removed;
- whether the calculation is a count, sum, mean, median, or percentage;
- the percentage denominator;
- the table relationship or join; and
- whether a game-level value was repeated across play rows.

Record one pattern you can state with certainty and one question that would
require more data or analysis.

*Your response:*

## Keep your work for reference

There is nothing to submit for Lab 2. Keep your Excel workbook, notes, and
rendered Quarto document for reference. Homework 2 requires similar analysis
with your selected dataset, so these files provide examples you can revisit.

### Completion checklist

- [ ] I identified each table's grain and key.
- [ ] I inspected and verified the join.
- [ ] I created a count and frequency table.
- [ ] I compared means and medians.
- [ ] I created and interpreted a cross-tabulation with percentages.
- [ ] I reproduced at least one result in Excel, Tableau when available, and R.
- [ ] I rendered `Lab_02_IP.pdf`.
