---
title: "Lab 5: Practice R Coding"
subtitle: "In-Person Sections - Student Version"
format:
  typst:
    toc: false
    papersize: us-letter
    fontsize: 10pt
    linestretch: 1
    margin:
      x: 1in
      y: 1in
    number-sections: false
    page-numbering: "1"
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(lubridate)
library(eco230r)
library(knitr)
```

## Purpose

Practice adapting working R examples. Complete at least 3 changes that help you
understand importing, selecting, transforming, summarizing, filtering, or
working with dates.

```{r}
airbnb <- read_csv("data/airbnb_chicago.csv", show_col_types = FALSE)
```

## 1) Select useful columns

Create `listing_data` containing `property_type`, `room_type`, `host_since`,
`number_of_reviews`, `beds`, `bedrooms`, `bathrooms`, `listed_price`, and
`review_scores_rating`.

```{r}

```

## 2) Calculate a field

Add `reviews_per_bed`. Decide how to handle listings with zero or missing beds.

```{r}

```

## 3) Summarize groups

Create a summary by `room_type` with a row count, median price, and mean review
score. Handle missing values explicitly.

```{r}

```

## 4) Filter rows

Keep listings with a finite `reviews_per_bed` value of at least 1.

```{r}

```

## 5) Create categories

Create a clearly labeled price category from `listed_price`. Explain why you
selected those breakpoints.

```{r}

```

*Your explanation:*

## 6) Parse a date

Convert `host_since` to a date and extract the host-start year.

```{r}

```

## 7) Combine several steps

Build one analysis that uses `select()`, `mutate()`, `filter()`, `group_by()`,
and `summarise()`. Print the result with `kable()`.

```{r}

```

## 8) Reflect

Which 3 examples did you alter successfully? Which 3 concepts were most
difficult or still need clarification?

*Your response:*

## Completion checklist

- [ ] I completed at least 3 meaningful code changes.
- [ ] I ran each completed chunk and examined its output.
- [ ] I explained any unresolved error or question.
- [ ] I rendered `Practice_R_Coding_IP.pdf`.
