---
title: "Understanding Data"
subtitle: "Start with the row, then understand the variables"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: false
execute:
  echo: true
  warning: false
  message: false
---

## The first question: what does one row represent?

Before calculating an average or making a graph, state the **unit of analysis**: the person, listing, transaction, day, crash, or other thing represented by one row.

If you misunderstand the row, every summary that follows can be wrong.

```{r}
#| label: load-sleep-data
#| echo: false
sleep <- read.csv(
  here::here("week01", "data", "sleep_tracking_Jan-Jun_2025_with_zip.csv"),
  check.names = FALSE
)

sleep$Date <- as.Date(sleep$Date)
knitr::kable(
  head(sleep[c("Date", "PersonName", "Location", "SleepEventType",
               "TotalMinutes", "SleepQuality")], 8),
  caption = "Eight rows from the ECO 230 sleep-tracking data"
)
```

In this file, one row is one recorded sleep event for one person. A person can have a main sleep event and a nap on the same date, so the grain is not simply "one person per day." If an analysis should include only main sleep, filter `SleepEventType == "Main"` before summarizing.

::: {.callout-important}
## Write the grain as a sentence

Use this pattern: **One row represents one [unit] for one [entity] during one [time period].**

Include any conditions created by filtering or aggregation.
:::

## Observations and variables

- An **observation** is one row.
- A **variable** is one column that records an attribute or measurement.
- An **identifier** distinguishes entities but is not automatically a quantity worth averaging.

| Variable | Role or type | Useful treatment |
|---|---|---|
| `Date` | Date/time | Sort, filter, group by week or month |
| `PersonName` | Nominal identifier | Group or count; do not average |
| `Location` | Nominal category | Frequencies, percentages, comparisons |
| `SleepEventType` | Nominal category | Filter or compare main sleep and naps |
| `TotalMinutes` | Quantitative, ratio | Mean, median, range, IQR, standard deviation |
| `SleepQuality` | Quantitative score | Compare distributions; interpret using its documented scale |
| `ZipCode` | Geographic identifier | Join or group; do not treat as a measured number |

An all-numeric column is not necessarily quantitative. ZIP codes, product codes, and patient IDs are labels even when stored as numbers.

## Data structure changes the question

### Cross-sectional data

A cross-section compares many entities at one point or period in time. For example, a single Airbnb snapshot can compare listings across neighborhoods.

### Time-series data

A time series follows one measure for one entity or aggregate over time. Daily average sleep duration for the entire sample is a time series.

### Panel data

Panel data follows multiple entities over time. The main-sleep records form a panel because the same people appear on many dates.

```{r}
#| label: panel-check
main_sleep <- sleep[sleep$SleepEventType == "Main", ]
person_days <- aggregate(Date ~ PersonName, main_sleep, function(x) length(unique(x)))
names(person_days)[2] <- "RecordedDays"
knitr::kable(person_days, caption = "Repeated observations make this a panel data set")
```

## Wide and long are descriptions, not judgments

The sleep-stage columns (`REMMinutes`, `DeepMinutes`, and `LightMinutes`) are stored wide: each stage has its own column. The same information could be made long with one `Stage` column and one `Minutes` column.

Wide data can be convenient for a report. Long data is often easier for grouped summaries and graphs. Choose the structure that supports the analysis, while keeping the grain explicit.

## A second example: Airbnb listings

```{r}
#| label: airbnb-example
#| echo: false
airbnb <- read.csv(
  here::here("shared", "data", "airbnb_chicago.csv"),
  check.names = FALSE
)

knitr::kable(
  head(airbnb[c("id", "room_type", "neighbourhood", "accommodates",
                 "listed_price", "review_scores_rating")], 6),
  digits = 1,
  caption = "Selected fields from the Chicago Airbnb data"
)
```

Here, one row represents one listing in the source snapshot. `listed_price` is a quantitative measure; `room_type` is categorical; `id` is an identifier. If you calculate one row per neighborhood, you have created a new, aggregated data set with a different grain.

## A five-question data check

Before analysis, answer:

1. What does one row represent?
2. What population and time period are covered?
3. Which columns are identifiers, categories, dates, and quantities?
4. Can the same entity appear more than once? If so, why?
5. What filtering, joining, or aggregation would change the grain?

::: {.callout-tip}
## A useful habit

Put the grain in a code comment immediately after loading data. When the grain changes, add another comment before the summary or join.
:::

## Background

This page replaces an earlier ECO 230 handout that used a CEO-compensation spreadsheet. The concepts and examples have been rewritten around current course data.
