---
title: "Homework 2: Exploratory Analysis"
subtitle: "In-Person Sections"
format:
  typst:
    toc: false
    papersize: us-letter
    fontsize: 10pt
    linestretch: 1
    margin:
      x: 1in
      y: 1in
    number-sections: false
    page-numbering: "1"
execute:
  echo: true
  warning: false
  message: false
editor: source
---

::: {.callout-note appearance="minimal" icon="false" title="Purpose"}
This assignment helps you describe the size, center, shape, and relationships in
one project dataset before creating charts or running hypothesis tests.

You will build four summary tables, explain the denominators and statistics you
selected, identify three findings, and reflect on responsible AI use.
:::

## Objectives

By the end of this assignment, you will be able to:

- create frequency, grouped-summary, and cross-tabulation tables;
- match measures of center and spread to the variable and distribution;
- distinguish row, column, and overall percentages;
- state denominators and missing-data rules; and
- interpret numerical summaries in context.

## Resources

- Week 2 slides on size, center, shape, relationships, and percentages
- Week 1 slides on grain, tidy data, dimensions, and measures
- The four `.csv` files in the `data` folder
- Excel, Tableau, or R in Posit Cloud

::: {.callout-tip appearance="minimal" icon="false" title="Expectations"}
This assignment is graded for clear, careful completion. Do not create charts.
Your tables must make their denominators, units, filters, and missing-data rules
clear.
:::

## 1) Choose and Load Your Dataset

Choose one project dataset. State its filename and any filters or cleaning steps
you apply.

**Dataset:**

**Filters or cleaning:**

If you are working in R, change `project_file` to the appropriate filename.

```{r}
project_file <- "accident_wi.csv"
df <- read.csv(file.path("data", project_file), check.names = FALSE)

cat(
  project_file, "contains", nrow(df), "rows and",
  ncol(df), "columns."
)
```

---

## 2) Select Your Tools

You may complete the tables in Excel, Tableau, R, or a combination. Insert each
finished table or its output in this document. Choose the tool that best supports
the table you are building.

**Tools used:**



---

## 3) Create Table 1 - Frequency Table

Create a frequency table for one categorical variable. A variable with at least
3-4 useful categories is recommended.

Your table must include:

- category;
- count (`n`);
- percent of total; and
- any other information needed to explain the distribution.

The following example shows the structure using Airbnb cancellation policy. Use
variables from your selected dataset in your final table.

```{r}
#| eval: false
example_df <- read.csv(file.path("data", "airbnb_chicago.csv"))

example_df |>
  count(cancellation_policy, name = "n") |>
  mutate(percent = n / sum(n)) |>
  arrange(desc(n))
```

**Table 1 code, table, or inserted output:**



**Brief finding:** What appears often, and what is rare?



---

## 4) Create Table 2 - Continuous by Categorical

Create a grouped summary table using:

- 1 continuous measure;
- 1 categorical dimension with at least 3 useful groups;
- `n_non_missing` for the measure; and
- a meaningful measure of center and spread.

Use mean and standard deviation when they represent the distribution well. Use
median and IQR when the distribution is skewed, contains influential outliers,
or is better described by its middle.

```{r}
#| eval: false
# Mean and standard-deviation structure:
df |>
  group_by(your_group_variable) |>
  summarise(
    n_non_missing = sum(!is.na(your_measure)),
    mean_value = mean(your_measure, na.rm = TRUE),
    sd_value = sd(your_measure, na.rm = TRUE)
  )

# Median and IQR structure:
df |>
  group_by(your_group_variable) |>
  summarise(
    n_non_missing = sum(!is.na(your_measure)),
    median_value = median(your_measure, na.rm = TRUE),
    q25 = quantile(your_measure, 0.25, na.rm = TRUE),
    q75 = quantile(your_measure, 0.75, na.rm = TRUE),
    iqr = q75 - q25
  )
```

**Table 2 code, table, or inserted output:**



In 1-2 sentences, explain why your measures of center and spread are meaningful.

*Your response:*



---

## 5) Create Table 3 - A Different Relationship

Create a second continuous-by-categorical summary that changes at least one of
these elements:

- the continuous measure;
- the categorical dimension;
- the summary choice; or
- the filter or analytical frame.

Include the same core fields as Table 2 and ensure that Table 3 answers a
different question.

```{r}
#| eval: false
# Add your Table 3 code here, or insert output from Excel or Tableau below.
```

**Table 3 code, table, or inserted output:**



**Brief finding:**



---

## 6) Create Table 4 - Cross-Tabulation

Create a cross-tabulation using two categorical variables, each with at least 3
useful categories.

Your table must include:

- cell counts;
- one clearly labeled percentage type: row, column, or overall;
- row and column totals; and
- a sentence that states the denominator.

```{r}
#| eval: false
xt <- table(df$your_row_variable, df$your_column_variable)
xt

# Choose one percentage calculation:
prop.table(xt, margin = 1)  # row percentages
prop.table(xt, margin = 2)  # column percentages
prop.table(xt)              # overall percentages
```

**Table 4 code, table, or inserted output:**



Complete this sentence: These percentages are calculated as __________, so each
__________ totals to 100%.

---

## Optional Correlation Table

If your dataset has two relevant continuous variables, you may report the number
of complete pairs, correlation `r`, and a one-sentence interpretation. This is
not required and does not replace Tables 1-4.

```{r}
#| eval: false
cor(df$your_numeric_variable_1, df$your_numeric_variable_2, use = "complete.obs")
```

---

## 7) Summarize Three Findings

Write 3 takeaways in approximately 5-10 sentences total. Each takeaway must:

- cite a specific table;
- include at least one value and its category or group; and
- explain what the value implies rather than only restating it.

**Example structure:** Table 2 shows that Group A has a higher median than Group
B, suggesting that...

*Your response:*



---

## 8) Reflect on Responsible AI Use

::: {.callout-warning appearance="minimal" icon="false" title="AI Policy"}
AI may support clarification, rewriting, or troubleshooting, but not statistical
selection or interpretation. Verify every claim and submit work you understand.
:::

Write 4-6 sentences explaining how you used AI, what it helped with, what it
missed or oversimplified, and one prompt you used. If you did not use AI, state
that directly.

*Your response:*



## Submission

Submit `Homework_02_IP.pdf` in Canvas. It must contain Tables 1-4, the three
findings, and the AI reflection. Also submit the supporting file you used to
create the tables when applicable: `.xlsx`, `.twbx`, or your Quarto source file.

## Checklist

- [ ] I identified my dataset, filters, tools, and missing-data decisions.
- [ ] I created four labeled tables and justified their summaries and denominators.
- [ ] I wrote three evidence-based findings and completed the AI reflection.
- [ ] I reviewed and submitted the PDF and required supporting files.

**Grading focus:** Your work will be evaluated for appropriate summaries,
correct denominators, readable tables, and interpretations tied to specific
evidence.
