---
title: "Lab 3: Visualizing Data with ggplot2"
subtitle: "R Companion - In-Person Sections"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-copy: true
    number-sections: false
execute:
  echo: true
  warning: false
  message: false
fig-width: 8
fig-height: 5
---

::: {.callout-note appearance="minimal" icon="false" title="Purpose"}
Use the same crash sample as the Tableau and Excel lab to make seven types of
visuals in R. Concentrate on the question, the summary, and the chart type.
The examples use a clean theme and a few deliberate design choices.

This is a worked companion, not an additional graded assignment. Keep following
the submission instructions for the main Lab 3 in Canvas.
:::

## Start here

1. Run `source("project_setup.R")` in the Console once to restore the packages.
2. Open `Lab_03_R_Companion_IP.qmd`. Run chunks from the top down, or click
   **Render** to create the complete HTML guide with code and charts.
3. Try one small change at a time. The **Try it** prompts are optional practice.

- [Open the Lab 3 project in Posit Cloud](https://posit.cloud/spaces/3173/content/12970507)
- [Watch the instructor's R walkthrough](https://mediaspace.wisconsin.edu/media/t/1_lqzvbpkk)
- [Read the walkthrough transcript](Lab_03_R_Transcript.md)

You will practice counting and averaging before plotting, choosing appropriate
geometries, and checking filters and percentage denominators. The data and its
provenance are included in `data/`; no download is needed to run the examples
after package setup. Use the Lab 3 video guide and hints in Canvas alongside
this document.

### Packages and a simple visual style

```{r}
#| label: packages-and-style
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(lubridate)
library(forcats)
library(stringr)

theme_set(
  theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          plot.title.position = "plot")
)

# A familiar Tableau-like palette; no extra theme package is needed.
tableau_colors <- c("#4E79A7", "#F28E2B", "#E15759", "#76B7B2",
                    "#59A14F", "#EDC948", "#B07AA1", "#FF9DA7",
                    "#9C755F", "#BAB0AC")
```

`theme_minimal()` supplies the overall style. We will also sort categories,
label units, and use color for groups when it helps answer the question.
Inside `aes()`, a column controls an appearance; outside `aes()`, a setting
such as `fill = tableau_colors[1]` gives every mark the same appearance.

### Read and check the data

```{r}
#| label: read-data
crashes <- read_csv("data/chicago_traffic_crashes_sample.csv",
                    show_col_types = FALSE) |>
  mutate(
    month = as.Date(floor_date(crash_datetime, "month")),
    quarter = factor(paste0("Q", quarter(crash_datetime)),
                     levels = c("Q1", "Q2", "Q3", "Q4"))
  )

nrow(crashes)
n_distinct(crashes$crash_id)
```

Both counts should be **12,000**. One row represents one sampled crash.
This deterministic classroom sample covers 2018-2025; its counts are not
official Chicago totals or crash rates. Weather measurements can be shared
by several crashes. Missing measurements stay missing, rather than becoming zero.

Each example starts from `crashes`, so a filter for one chart does not silently
carry into another. `|>` passes a result to the next step; `+` adds a plot layer.

## 1) Ranking: count, sort, then plot

**Question:** Which reported causes appear most often?

Exclude missing causes, UNABLE TO DETERMINE, and NOT APPLICABLE before
selecting the ten largest groups, matching the lab's ranking example.

```{r}
#| label: ranking
ranking <- crashes |>
  filter(!is.na(primary_contributory_cause),
         !primary_contributory_cause %in%
           c("UNABLE TO DETERMINE", "NOT APPLICABLE")) |>
  count(primary_contributory_cause, name = "crash_count") |>
  slice_max(crash_count, n = 10, with_ties = FALSE)

ranking_plot <- ggplot(ranking,
                      aes(x = crash_count,
                          y = fct_reorder(primary_contributory_cause,
                                          crash_count))) +
  geom_col(fill = tableau_colors[1]) +
  scale_y_discrete(labels = function(x) str_wrap(str_to_sentence(x), 32)) +
  scale_x_continuous(labels = scales::label_comma()) +
  labs(title = "Which reported causes appear most often?",
       subtitle = "Ten most common causes after the stated exclusions",
       x = "Sampled crashes", y = NULL)

ranking_plot
```

`count()` makes one row per cause. `geom_col()` uses those already-calculated
counts as bar lengths. `fct_reorder()` puts the largest bar at the top.

**Check:** There are ten bars. Each bar summarizes crashes, rather than one crash.

**Try it:** Change `n = 10` to `n = 5`.

## 2) Nominal comparison: averages need group sizes

**Question:** How does average injury count differ across causes?

Start again from all crashes. Keep causes with at least 30 crashes and exclude
NOT APPLICABLE and missing causes. UNABLE TO DETERMINE is retained here, as in
the lab after removing the ranking selection.

```{r}
#| label: average-injuries
#| fig-height: 10
injury_summary <- crashes |>
  filter(!is.na(primary_contributory_cause),
         primary_contributory_cause != "NOT APPLICABLE") |>
  group_by(primary_contributory_cause) |>
  summarise(crash_count = n(),
            measured_count = sum(!is.na(injuries_total)),
            mean_injuries = mean(injuries_total, na.rm = TRUE),
            .groups = "drop") |>
  filter(crash_count >= 30, measured_count > 0)

injury_plot <- ggplot(injury_summary,
                     aes(x = mean_injuries,
                         y = fct_reorder(primary_contributory_cause,
                                         mean_injuries))) +
  geom_col(fill = tableau_colors[2]) +
  scale_y_discrete(labels = function(x) str_wrap(str_to_sentence(x), 34)) +
  theme(axis.text.y = element_text(size = 10)) +
  labs(title = "How does average injury count differ across causes?",
       subtitle = "At least 30 sampled crashes per cause",
       x = "Average injuries per crash with a recorded injury count", y = NULL)

injury_plot
```

Unlike a Tableau tooltip, a static chart does not reveal the group count on
hover. Inspect the supporting table:

```{r}
#| label: inspect-group-sizes
injury_summary |>
  arrange(desc(mean_injuries)) |>
  select(primary_contributory_cause, crash_count, measured_count, mean_injuries) |>
  knitr::kable(digits = 2)
```

**Check:** An average is not a total. The 30-crash cutoff does not make every
group equally reliable. `measured_count` is the denominator of each average;
missing injury counts are omitted from that calculation.

**Try it:** Raise the cutoff to 100 and see which groups remain.

## 3) Time series: keep month and year together

**Question:** How do monthly counts differ across reported weather conditions?

Exclude CLEAR and UNKNOWN for this version. Small multiples keep the remaining
conditions readable without a crowded legend. Complete missing month-condition
combinations with zero because the complete sample contains no crashes in those
combinations; this is not a rule for filling missing weather measurements.

```{r}
#| label: monthly-counts
#| fig-height: 7
months <- seq(min(crashes$month), max(crashes$month), by = "month")

monthly <- crashes |>
  filter(!is.na(reported_weather_condition),
         !reported_weather_condition %in% c("CLEAR", "UNKNOWN")) |>
  count(month, reported_weather_condition, name = "crash_count") |>
  complete(month = months, reported_weather_condition,
           fill = list(crash_count = 0))

time_plot <- ggplot(monthly, aes(x = month, y = crash_count)) +
  geom_line(color = tableau_colors[1], linewidth = 0.5) +
  facet_wrap(~ reported_weather_condition, ncol = 3,
             scales = "free_y", labeller = label_wrap_gen(20)) +
  scale_x_date(date_breaks = "4 years", date_labels = "%Y") +
  labs(title = "How do monthly counts differ by weather?",
       subtitle = "CLEAR and UNKNOWN excluded; each panel has its own vertical scale",
       x = NULL, y = "Sampled crashes per month")

time_plot
```

**Check:** January in one year stays separate from January in another year.
Compare timing and shape within panels; their heights do not represent the same
counts when `scales = "free_y"`.

**Try it:** Change the facet setting to `scales = "fixed"` to compare magnitudes.

## 4) Part-to-whole: make the denominator explicit

**Question:** Within each quarter, what share belongs to each injury category?

These Q1-Q4 groups pool the same quarter across years, matching the lab.
Exclude missing injury categories, then divide by all remaining crashes in
each quarter.

```{r}
#| label: quarterly-shares
quarter_shares <- crashes |>
  filter(!is.na(most_severe_injury)) |>
  count(quarter, most_severe_injury, name = "crash_count") |>
  group_by(quarter) |>
  mutate(share = crash_count / sum(crash_count)) |>
  ungroup()

quarter_shares |>
  group_by(quarter) |>
  summarise(total_share = sum(share), .groups = "drop")

injury_colors <- c("FATAL" = "#E15759",
                   "INCAPACITATING INJURY" = "#F28E2B",
                   "NONINCAPACITATING INJURY" = "#4E79A7",
                   "REPORTED, NOT EVIDENT" = "#EDC948",
                   "NO INDICATION OF INJURY" = "#BAB0AC")

share_plot <- ggplot(quarter_shares,
                     aes(x = quarter, y = share, fill = most_severe_injury)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = injury_colors,
                    labels = function(x) str_wrap(str_to_sentence(x), 24)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(title = "Within each quarter, how are injury categories divided?",
       subtitle = "Q1-Q4 pooled across years; missing injury categories excluded",
       x = NULL, y = "Share of sampled crashes in the quarter", fill = NULL)

share_plot
```

Every `total_share` should equal 1 (100%). Now remove two categories from the
display **after** calculating the shares:

```{r}
#| label: hidden-categories
visible_shares <- quarter_shares |>
  filter(!most_severe_injury %in%
           c("NO INDICATION OF INJURY", "REPORTED, NOT EVIDENT"))

hidden_plot <- ggplot(visible_shares,
                      aes(x = quarter, y = share, fill = most_severe_injury)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = injury_colors,
                    labels = function(x) str_wrap(str_to_sentence(x), 24)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(title = "What share remains visible after hiding two categories?",
       subtitle = "Denominator: all nonmissing injury categories within each quarter",
       x = NULL, y = "Share of sampled crashes in the quarter", fill = NULL)

hidden_plot
```

The named colors keep each injury category the same color in both charts. The
visible stacks no longer reach 100%. Filtering the raw crashes **before** calculating
shares would instead change the denominator. Do not use
`position = "fill"` on this reduced table: it would rescale the visible groups
back to 100%.

**Try it:** Inspect `visible_shares` and explain the denominator of one row.

## 5) Relationship: one point per crash

**Question:** How are temperature and dew point related?

```{r}
#| label: weather-relationship
weather_pairs <- crashes |>
  filter(!is.na(weather_station_name),
         !is.na(temperature_f), !is.na(dew_point_f))

relationship_plot <- ggplot(weather_pairs,
                            aes(x = temperature_f, y = dew_point_f)) +
  geom_point(color = tableau_colors[1], alpha = 0.12, size = 0.8) +
  facet_wrap(~ weather_station_name, labeller = label_wrap_gen(22)) +
  labs(title = "How are temperature and dew point related?",
       x = "Temperature (degrees F)", y = "Dew point (degrees F)")

relationship_plot
```

**Check:** No `summarise()` occurs before plotting, so each point represents a
crash with complete paired measurements and a station name. Repeated station-hour
measurements can put several crashes at the same position. Transparency helps
show overlap; these are not independent weather observations.

**Try it:** Change `alpha` from 0.12 to 0.5 and compare how well you can see overlap.

## 6) Distribution: bins and box plots

**Question:** What temperatures occur in the sample, and how do they differ by
reported weather condition?

```{r}
#| label: temperature-histogram
temperatures <- crashes |>
  filter(!is.na(temperature_f))

histogram_plot <- ggplot(temperatures, aes(x = temperature_f)) +
  geom_histogram(binwidth = 5, boundary = 0,
                 fill = tableau_colors[1], color = "white") +
  labs(title = "What temperatures occur in the sampled crashes?",
       subtitle = "5-degree bins; crashes with missing temperature omitted",
       x = "Temperature (degrees F)", y = "Sampled crashes")

histogram_plot
```

Now summarize the individual temperatures within each weather category:

```{r}
#| label: temperature-boxplots
box_data <- temperatures |>
  filter(!is.na(reported_weather_condition))

box_plot <- ggplot(box_data,
                   aes(x = temperature_f,
                       y = fct_reorder(reported_weather_condition,
                                       temperature_f, .fun = median))) +
  geom_boxplot(fill = tableau_colors[4], outlier.alpha = 0.3) +
  scale_y_discrete(labels = function(x) str_to_sentence(x)) +
  labs(title = "How do temperatures differ by reported weather?",
       x = "Temperature (degrees F)", y = NULL)

box_plot
```

**Check:** A box shows the middle 50% of observations and a line at the median.
Points beyond the whiskers are unusual relative to that group, not necessarily
errors. Small weather groups deserve caution. Both plots describe temperatures
attached to sampled crashes, not the distribution of all Chicago weather.

**Try it:** Compare histogram bin widths of 2, 5, and 10 degrees F.

## 7) Point map: coordinates without a map service

**Question:** Where are the sampled crashes located?

This simple coordinate map needs no basemap download, spatial package, or API
key. Missing coordinates are omitted. To keep the legend manageable, retain the
four most frequent trafficway labels among mapped crashes and combine the rest.

```{r}
#| label: crash-map
#| fig-height: 7
map_data <- crashes |>
  filter(!is.na(longitude), !is.na(latitude)) |>
  mutate(trafficway_group = fct_lump_n(trafficway_type, n = 4,
                                      other_level = "ALL OTHER TYPES"))

nrow(crashes) - nrow(map_data) # Crashes omitted because coordinates are missing

map_plot <- ggplot(map_data,
                   aes(x = longitude, y = latitude, color = trafficway_group)) +
  geom_point(alpha = 0.25, size = 0.6) +
  coord_fixed(ratio = 1.34) +
  scale_color_manual(values = tableau_colors,
                     labels = function(x) str_wrap(str_to_sentence(x), 24)) +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 3))) +
  labs(title = "Where are the sampled crashes located?",
       subtitle = "One point per crash with coordinates; no basemap",
       x = "Longitude (degrees)", y = "Latitude (degrees)", color = "Trafficway")

map_plot
```

`coord_fixed(ratio = 1.34)` approximately adjusts for longitude/latitude spacing
near Chicago. This is a local display adjustment, not a projected map for measuring
distances. Overlapping points can hide crashes. More sampled crashes in an area
does not establish greater risk per traveler.

**Try it:** Color by `reported_weather_condition` instead. Decide whether the
additional categories help answer a clear question.

## Save a chart and check your work

Run this chunk to export the ranking chart as a PNG you can place in Word:

```{r}
#| label: save-chart
#| eval: false
ggsave("lab3-ranking.png", plot = ranking_plot,
       width = 8, height = 5, dpi = 300)
```

Before reusing a chart, identify what one mark represents, check the aggregation,
state exclusions, and name the denominator of any percentage. Explain one design
choice that makes the comparison easier to read. There is no additional upload
for this companion unless your instructor requests one.

### Function references

- [ggplot2 themes](https://ggplot2.tidyverse.org/reference/ggtheme.html)
- [Bar charts with precomputed values](https://ggplot2.tidyverse.org/reference/geom_bar.html)
- [Small multiples with facets](https://ggplot2.tidyverse.org/reference/facet_wrap.html)
- [Fixed coordinate aspect ratios](https://ggplot2.tidyverse.org/reference/coord_fixed.html)
