---
title: "Week 8 - Comparing Two Groups"
format:
  revealjs:
    theme:
      - default
      - ../shared/styles/boland-reveal.scss
    slide-number: true
    hash: true
    center: false
    transition: fade
    controls: true
    pdf-export: true
execute:
  echo: false
---

## Steps in Hypothesis Testing

1.  [*What statistical story am I attempting to tell?*]{.muted}

::: {.callout-tip title="What are we trying to find out?"}
:::

2.  [*What have I estimated or plotted... If no uncertainty, what should the business do?*]{.muted}

::: {.callout-tip title="What do we expect to happen?"}
:::

3.  [*Which test is appropriate and what is the null hypothesis?*]{.muted}

::: {.callout-tip title="How are we going to test it?"}
:::

4.  [*What is the statistical interpretation of the test result?*]{.muted}

::: {.callout-tip title="Do we believe it?"}
:::

5.  [*What is the practical interpretation of this analysis for a business audience?*]{.muted}

::: {.callout-tip title="Does it matter?"}
:::

# t-Tests

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(eco230r)
library(here)
library(readr)

data("t_tall", package = "eco230r")
data("t_wide", package = "eco230r")
data("tea_tab", package = "eco230r")
data("napData", package = "eco230r")

```

## Independent Samples t-Test `idt()`

```{r run_idt echo:FALSE, include:FALSE, message=FALSE, warning=FALSE, results="hide"}
df <- t_tall

h3 <- df %>%
  idt(scones ~ tea)
```

:::::::::: columns
::::: {.column width="40%"}
<br>

::: fragment
### Steps 1-3: Setup

<br>

**Test:** Independent Samples t-test\
**Function:** `idt()`

<br>

**H**<sub>0</sub>: μ<sub>Black</sub> = μ<sub>Green</sub>\
**H**<sub>A</sub>: μ<sub>Black</sub> $\neq$ μ<sub>Green</sub>\
**α = 0.05**
:::

<br>

::: fragment
### Run the Test

```{r show_idt, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
df %>%
  idt(scones ~ tea)
```
:::
:::::

::: {.column width="5%"}
:::

::::: {.column width="55%"}
<br>

::: fragment
### Step 4: Do we believe it?

`h3$results`

`r h3$results`

<br>

**Decision Rule:**

-   If p $\leq \alpha$ → Reject H<sub>0</sub>\
-   If p \> $\alpha$ → Fail to Reject H<sub>0</sub>

<br>
:::

::: fragment
**Decision:**\
p = 0.107 \> 0.05 → **Fail to Reject H**<sub>0</sub>
:::
:::::
::::::::::

## Bayes Factor vs. Effect Size `idt()` \| `pst()` \| `ost()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
**`r h3$results`**
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

:::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 185px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

::: {style="text-align:center; font-size:1.02rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; align-items:center; justify-content:center;"}
Cohen's $d$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
**0.33 – 1**
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
**1 – 3**
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
**3 – 10**
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem; margin-top:.2rem;"}
$d=.80$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Looks large, but evidence favors** $H_0$\
Possible noisy result or very small sample
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Large-looking difference, weak evidence**\
Promising, but still uncertain
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Potentially important effect**\
Need more evidence before being confident
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and practically meaningful**\
A noticeable difference
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Big and well supported**\
Strong evidence for a substantial effect
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem; margin-top:.2rem;"}
$d=.50$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate observed difference, but evidence favors** $H_0$\
Could be unstable
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate-looking effect, weak evidence**\
Unclear whether it will replicate
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Some indication of a meaningful effect**\
Not decisive yet
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and noticeable**\
Worth taking seriously
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Meaningful effect with strong support**\
Good evidence and practical relevance
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem; margin-top:.2rem;"}
$d=.20$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Little effect, evidence favors** $H_0$\
Probably not much happening
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Small difference, weak evidence overall**\
Not much to conclude
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Maybe real, but still small**\
Even if true, impact is modest
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real, but modest**\
Useful distinction: real does not mean big
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Very strong evidence for a small effect**\
Consistent, but not large
:::
::::::::::::::::::::::::::::::

## Step 5: Does it matter?

::::: columns
::: {.column width="60%"}
<br>

**Statistical:**

-   No statistically significant difference between groups

**Practical:**

-   Effect size: $d$ = -0.686 → *medium effect*\
-   Bayes Factor: bf<sub>10</sub> = 1.013 → *anecdotal evidence for H*<sub>A</sub>

**Takeaway:**

-   Evidence is weak\
-   Data does not convincingly support a difference\
-   Any observed difference may be noise
:::

::: {.column width="40%"}
<br>

`h3$descriptive_statistics`

```{r idt_desc echo:TRUE, include:TRUE}
h3$descriptive_statistics
```
:::
:::::

## Paired Samples t-Test `pst()`

```{r run_pst echo:FALSE, include:FALSE, message=FALSE, warning=FALSE, results="hide"}
df <- t_wide

h4 <- df %>%
  pst(~black_tea,~green_tea)
```

:::::::::: columns
::::: {.column width="40%"}
<br>

::: fragment
### Steps 1-3: Setup

<br>

**Test:** Paired Samples t-test\
**Function:** `pst()`

<br>

**H**<sub>0</sub>: μ<sub>Black</sub> = μ<sub>Green</sub>\
**H**<sub>A</sub>: μ<sub>Black</sub> $\neq$ μ<sub>Green</sub>\
**α = 0.05**
:::

<br>

::: fragment
### Run the Test

```{r show_pst, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
df %>%
  pst(~black_tea,~green_tea)
```
:::
:::::

::: {.column width="5%"}
:::

::::: {.column width="55%"}
<br>

::: fragment
### Step 4: Do we believe it?

`h4$results`

`r h4$results`

<br>

**Decision Rule:**

-   If p $\leq \alpha$ → Reject H<sub>0</sub>\
-   If p \> $\alpha$ → Fail to Reject H<sub>0</sub>

<br>
:::

::: fragment
**Decision:**\
p = 0.031 \< 0.05 → **Reject H**<sub>0</sub>
:::
:::::
::::::::::

## Bayes Factor vs. Effect Size `idt()` \| `pst()` \| `ost()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
**`r h4$results`**
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

:::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 185px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

::: {style="text-align:center; font-size:1.02rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; align-items:center; justify-content:center;"}
Cohen's $d$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
**0.33 – 1**
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
**1 – 3**
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
**3 – 10**
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem; margin-top:.2rem;"}
$d=.80$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Looks large, but evidence favors** $H_0$\
Possible noisy result or very small sample
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Large-looking difference, weak evidence**\
Promising, but still uncertain
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Potentially important effect**\
Need more evidence before being confident
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and practically meaningful**\
A noticeable difference
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Big and well supported**\
Strong evidence for a substantial effect
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem; margin-top:.2rem;"}
$d=.50$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate observed difference, but evidence favors** $H_0$\
Could be unstable
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate-looking effect, weak evidence**\
Unclear whether it will replicate
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Some indication of a meaningful effect**\
Not decisive yet
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and noticeable**\
Worth taking seriously
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Meaningful effect with strong support**\
Good evidence and practical relevance
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem; margin-top:.2rem;"}
$d=.20$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Little effect, evidence favors** $H_0$\
Probably not much happening
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Small difference, weak evidence overall**\
Not much to conclude
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Maybe real, but still small**\
Even if true, impact is modest
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real, but modest**\
Useful distinction: real does not mean big
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Very strong evidence for a small effect**\
Consistent, but not large
:::
::::::::::::::::::::::::::::::

## Step 5: Does it matter?

::::: columns
::: {.column width="60%"}
**Statistical:**

-   Statistically significant difference between groups

<br>

**Practical:**

-   Effect size: $d$ = -0.714 → *medium to large effect*\
-   Bayes Factor: bf<sub>10</sub> = 1.013 → *anecdotal evidence for H*<sub>A</sub>

<br>

**Takeaway:**

-   Evidence is weak\
-   Some indication of a meaningful effect
-   Need more evidence before being confident
:::

::: {.column width="40%"}
<br>

`h4$descriptive_statistics`

```{r pst_desc echo:TRUE, include:TRUE}
h4$descriptive_statistics
```
:::
:::::

## Wilcoxon Rank-Sum Test `idw()`

```{r run_idw echo:FALSE, include:FALSE, message=FALSE, warning=FALSE, results="hide"}
df <- t_tall

h5 <- df %>%
  idw(scones~tea)
```

:::::::::: columns
::::: {.column width="40%"}
<br>

::: fragment
### Steps 1-3: Setup

<br>

**Test:** Wilcoxon Rank-Sum Test\
**Function:** `idw()`

<br>

**H**<sub>0</sub>: Median<sub>Black</sub> = Median<sub>Green</sub>\
**H**<sub>A</sub>: Median<sub>Black</sub> $\neq$ Median<sub>Green</sub>\
**α = 0.05**
:::

<br>

::: fragment
### Run the Test

```{r show_idw, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
df %>%
  idw(scones~tea)
```
:::
:::::

::: {.column width="5%"}
:::

::::: {.column width="55%"}
<br>

::: fragment
### Step 4: Do we believe it?

`r h5$results`

<br>

**Decision Rule:**

-   If p $\leq \alpha$ → Reject H<sub>0</sub>\
-   If p \> $\alpha$ → Fail to Reject H<sub>0</sub>

<br>
:::

::: fragment
**Decision:**\
p = 0.137 \> 0.05 → **Fail to Reject H**<sub>0</sub>
:::
:::::
::::::::::

## Bayes Factor vs. Effect Size `idw()` \| `psw()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
**`r h3$results`**
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

:::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 185px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

::: {style="text-align:center; font-size:1.02rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; align-items:center; justify-content:center;"}
Rank-biserial $r_{rb}$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
0.33–1
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
1–3
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
3–10
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem; margin-top:.2rem;"}
$r_{rb}=.50$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Strong rank separation, but evidence favors** $H_0$\
Possible noisy result or very small sample
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Strong-looking separation, weak evidence**\
Promising, but still uncertain
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Potentially important separation**\
Need more evidence before being confident
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and practically meaningful**\
Ranks are noticeably separated
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Strong and well supported**\
Clear evidence of substantial separation
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem; margin-top:.2rem;"}
$r_{rb}=.30$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate rank separation, but evidence favors** $H_0$\
Could be unstable
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate-looking separation, weak evidence**\
Unclear whether it will replicate
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Some indication of meaningful separation**\
Not decisive yet
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and noticeable**\
Worth taking seriously
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Meaningful and well supported**\
Good evidence and practical relevance
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem; margin-top:.2rem;"}
$r_{rb}=.10$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Little rank separation, evidence favors** $H_0$\
Probably not much happening
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Small separation, weak evidence overall**\
Not much to conclude
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Maybe real, but still small**\
Even if true, impact is modest
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real, but modest**\
Useful distinction: real does not mean large
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Very strong evidence for a small effect**\
Consistent, but not large
:::
::::::::::::::::::::::::::::::

## Bayes Factor vs. Effect Size `csf()` \| `csi()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
r h1\$results
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

::::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 205px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

:::: {style="text-align:center; font-size:.96rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; align-items:center; justify-content:center; line-height:1.15;"}
::: {style="font-size:.82rem; font-weight:600; margin-top:.12rem;"}
Cohen's $w$ (Association) Cramer's $V$ (Departure Size)
:::
::::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
0.33–1
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
1–3
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
3–10
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .50$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Large pattern, but evidence favors** $H_0$\
Possible noisy result or very small sample
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Large-looking pattern, weak evidence**\
Promising, but still uncertain
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Potentially important association / mismatch**\
Need more evidence before being confident
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and practically meaningful**\
A strong pattern in the counts
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Large and well supported**\
Strong evidence for a substantial pattern
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .30$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate pattern, but evidence favors** $H_0$\
Could be unstable
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate-looking pattern, weak evidence**\
Unclear whether it will hold up
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Some indication of a meaningful pattern**\
Not decisive yet
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and noticeable**\
Worth taking seriously
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Meaningful and well supported**\
Good evidence and practical relevance
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .10$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Little pattern, evidence favors** $H_0$\
Probably not much happening
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Small pattern, weak evidence overall**\
Not much to conclude
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Maybe real, but still small**\
Even if true, impact is modest
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real, but modest**\
Useful distinction: real does not mean large
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Very strong evidence for a small pattern**\
Consistent, but not large
:::
:::::::::::::::::::::::::::::::

## Bayes Factor vs. Effect Size `ano()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
r h1\$results
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

::::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 205px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

:::: {style="text-align:center; font-size:.96rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; flex-direction:column; align-items:center; justify-content:center; line-height:1.1;"}
Variance Explained

::: {style="font-size:.82rem; font-weight:600;"}
$\omega^2$, $\xi$
:::
::::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
0.33–1
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
1–3
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
3–10
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem;"}
$\omega^2, \xi = .14$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Large variance explained, but evidence favors** $H_0$\
Group differences appear large, but this may be due to noise or small sample size
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Large-looking differences, weak evidence**\
Groups look different, but we are not yet confident this pattern is real
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Potentially important group differences**\
Group membership may explain a meaningful portion of the outcome, but evidence is still limited
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real and meaningful differences**\
Group membership explains a substantial portion of variation in the outcome
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Large and well supported**\
Strong evidence that group differences explain a large share of the outcome
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem;"}
$\omega^2, \xi = .06$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Moderate variance explained, but evidence favors** $H_0$\
There may be some group differences, but the data do not support them clearly
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Moderate differences, weak evidence**\
Groups show some separation, but it may not hold up with more data
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Some indication of meaningful differences**\
Group membership may explain part of the outcome, but not decisively
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real and noticeable differences**\
Group differences explain a meaningful portion of variation
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Meaningful and well supported**\
Good evidence that group differences matter for the outcome
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem;"}
$\omega^2, \xi = .01$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Little variance explained, evidence favors** $H_0$\
Group differences do not explain much of the outcome
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Small differences, weak evidence**\
Groups look similar, and there is little evidence of meaningful separation
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Maybe real, but small impact**\
Even if group differences exist, they explain very little of the outcome
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real, but modest differences**\
Group differences are consistent, but explain only a small portion of variation
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Very strong evidence for small differences**\
Group differences are real, but they explain very little of the outcome
:::
:::::::::::::::::::::::::::::::

## Bayes Factor vs. Effect Size `slr()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
r h1\$results
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

::::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 205px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

:::: {style="text-align:center; font-size:.96rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; flex-direction:column; align-items:center; justify-content:center; line-height:1.1;"}
Variance Explained

::: {style="font-size:.82rem; font-weight:600;"}
$R^2$
:::
::::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
0.33–1
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
1–3
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
3–10
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem;"}
$R^2 = .26$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Large variance explained, but evidence favors** $H_0$\
The predictor appears important, but this may be due to noise or limited data
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Strong-looking relationship, weak evidence**\
The predictor explains a lot of variation, but we are not yet confident it is real
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Potentially important predictor**\
The predictor may explain a large portion of the outcome, but evidence is still limited
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real and meaningful relationship**\
The predictor explains a substantial share of variation in the outcome
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Large and well supported**\
Strong evidence that the predictor is highly informative for the outcome
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem;"}
$R^2 = .13$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Moderate variance explained, but evidence favors** $H_0$\
The predictor may matter, but the data do not support it clearly
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Moderate relationship, weak evidence**\
The predictor explains some variation, but this may not hold up
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Some indication of a useful predictor**\
The predictor may explain part of the outcome, but not decisively
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real and noticeable relationship**\
The predictor explains a meaningful portion of variation
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Meaningful and well supported**\
Good evidence that the predictor contributes to the outcome
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem;"}
$R^2 = .02$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Little variance explained, evidence favors** $H_0$\
The predictor does not explain much of the outcome
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Small relationship, weak evidence**\
The predictor explains very little variation
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Maybe real, but small impact**\
Even if real, the predictor has limited influence on the outcome
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real, but modest relationship**\
The predictor consistently explains a small portion of variation
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Very strong evidence for a small relationship**\
The predictor matters, but only explains a small share of the outcome
:::
:::::::::::::::::::::::::::::::
