---
title: "Practice Practicum"
subtitle: "In-Person Sections"
format:
  typst:
    toc: false
    papersize: us-letter
    fontsize: 10pt
    linestretch: 1
    margin:
      x: 1in
      y: 1in
    number-sections: false
    page-numbering: "1"
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(eco230r)
library(knitr)
```

## Purpose

Use this practice project to rehearse the applied-analysis portion of the final.
Read `prompt/Practice_Practicum_Prompt.qmd` and
`prompt/Practice_Practicum_Guidelines.qmd` before beginning.

## Load the dataset

```{r}
student_athletes <- read_csv(
  "data/studentathletes.csv",
  show_col_types = FALSE
)

glimpse(student_athletes)
```

State the unit of analysis and identify which fields address the question you
selected from the prompt.

*Your response:*

## Analysis

Build at least one appropriate table, one graph, and one hypothesis test. Keep
the analysis focused on one question if that produces a clearer result.

```{r}

```

## Findings and limitations

Summarize the most useful findings and important limitations in concise bullet
points for the decision maker described in the prompt.

- Finding:
- Finding:
- Limitation:

## Completion checklist

- [ ] I used at least two required software tools.
- [ ] I created a table, graph, and appropriate hypothesis test.
- [ ] I interpreted the result in context.
- [ ] I identified at least one limitation.
- [ ] I rendered `Practicum_Practice_IP.pdf`.
