---
title: "Interpreting Chi-Square Analysis"
format:
  revealjs:
    theme:
      - default
      - ../shared/styles/boland-reveal.scss
    slide-number: true
    hash: true
    center: false
    transition: fade
    controls: true
    pdf-export: true
execute:
  echo: false
---

## Steps in Hypothesis Testing

1.  [*What statistical story am I attempting to tell?*]{.muted}

::: {.callout-tip title="What are we trying to find out?"}
:::

2.  [*What have I estimated or plotted... If no uncertainty, what should the business do?*]{.muted}

::: {.callout-tip title="What do we expect to happen?"}
:::

3.  [*Which test is appropriate and what is the null hypothesis?*]{.muted}

::: {.callout-tip title="How are we going to test it?"}
:::

4.  [*What is the statistical interpretation of the test result?*]{.muted}

::: {.callout-tip title="Do we believe it?"}
:::

5.  [*What is the practical interpretation of this analysis for a business audience?*]{.muted}

::: {.callout-tip title="Does it matter?"}
:::

# Chi-Square Analysis

Comparing Categorical Variables

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(eco230r)
library(here)
library(readr)

load(system.file("data", "t_tall.rda", package = "eco230r"))
load(system.file("data", "t_wide.rda", package = "eco230r"))
load(system.file("data", "tea_tab.rda", package = "eco230r"))
load(system.file("data", "napData.rda", package = "eco230r"))

```

## Chi-Square Goodness of Fit `csf()`

```{r run_csf echo:FALSE, include:FALSE, message=FALSE, warning=FALSE, results="hide"}
df <- tea_tab

h1 <- df %>%
  csf(~cool_time)
```

:::::::::: columns
::::: {.column width="40%"}
<br>

::: fragment
### Steps 1-3: Setup

<br>

**Test:** Chi-Square Goodness of Fit Test\

**Function:** `csf()`

<br>

**H<sub>0</sub>:** Cool_Time fits expected model

**H<sub>A</sub>:** Cool_Time fits expected model

**α =** 0.05
:::

<br>

::: fragment
### Run the Test

```{r show_idt, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
h1 <- df %>%
  csf(~cool_time)
```
:::
:::::

::: {.column width="5%"}
:::

::::: {.column width="55%"}
<br>

::: fragment
### Step 4: Do we believe it?

`h1$results`

`r h1$results`

<br>

**Decision Rule:**

-   If p $\leq \alpha$ → Reject H<sub>0</sub>\
-   If p \> $\alpha$ → Fail to Reject H<sub>0</sub>

<br>
:::

::: fragment
**Decision:**\
p = 0.228 \> 0.05 → **Fail to Reject H<sub>0</sub>**
:::
:::::
::::::::::

## Bayes Factor vs. Effect Size `csf()` \| `csi()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
**`r h1$results`**
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

::::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 205px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

:::: {style="text-align:center; font-size:.96rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; align-items:center; justify-content:center; line-height:1.15;"}
::: {style="font-size:.82rem; font-weight:600; margin-top:.12rem;"}
Cohen's $w$ (Association) Cramer's $V$ (Departure Size)
:::
::::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
0.33–1
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
1–3
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
3–10
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .50$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Large pattern, but evidence favors** $H_0$\
Possible noisy result or very small sample
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Large-looking pattern, weak evidence**\
Promising, but still uncertain
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Potentially important association / mismatch**\
Need more evidence before being confident
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and practically meaningful**\
A strong pattern in the counts
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Large and well supported**\
Strong evidence for a substantial pattern
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .30$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate pattern, but evidence favors** $H_0$\
Could be unstable
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate-looking pattern, weak evidence**\
Unclear whether it will hold up
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Some indication of a meaningful pattern**\
Not decisive yet
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and noticeable**\
Worth taking seriously
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Meaningful and well supported**\
Good evidence and practical relevance
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .10$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Little pattern, evidence favors** $H_0$\
Probably not much happening
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Small pattern, weak evidence overall**\
Not much to conclude
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Maybe real, but still small**\
Even if true, impact is modest
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real, but modest**\
Useful distinction: real does not mean large
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Very strong evidence for a small pattern**\
Consistent, but not large
:::
:::::::::::::::::::::::::::::::

## Step 5: Does it matter?

::::: columns
::: {.column width="60%"}
<br>

**Statistical:**

-   Variable is not significantly different than the expected model where counts from all categories are equal

**Practical:**

-   Effect size: $w$ = 0.251 → *small to medium effect*\
-   Bayes Factor: bf<sub>10</sub> = 0.082 → *strong evidence for H<sub>0</sub>*

**Takeaway:**

-   Little to moderate pattern in data, evidence favors H<sub>0</sub>
-   Pattern could be unstable
-   Possible sample is too small to see meaningful pattern
:::

::: {.column width="40%"}
<br>

`h1$observed`

```{r csf_obs echo:TRUE, include:TRUE}
h1$observed
```

`h1$expected`

```{r csf_exp echo:TRUE, include:TRUE}
h1$expected
```

`h1$standardized_residuals`

```{r csf_sr echo:TRUE, include:TRUE}
h1$standardized_residuals
```
:::
:::::

## Chi-Square Test of Association `csi()`

```{r run_csi echo:FALSE, include:FALSE, message=FALSE, warning=FALSE, results="hide"}
df <- tea_tab

h2 <- df %>%
  csi(finger_preference~cool_time)
```

:::::::::: columns
::::: {.column width="40%"}
<br>

::: fragment
### Steps 1-3: Setup

<br>

**Test:** Chi-Square Test of Association\

**Function:** `csi()`

<br>

**H<sub>0</sub>:** Finger_Preference not associated with Cool_Time

**H<sub>A</sub>:** Finger_Preference is associated with Cool_Time

**α =** 0.05
:::

<br>

::: fragment
### Run the Test

```{r show_csi, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
h2 <- df %>%
  csi(finger_preference~cool_time)
```
:::
:::::

::: {.column width="5%"}
:::

::::: {.column width="55%"}
<br>

::: fragment
### Step 4: Do we believe it?

`h2$results`

`r h2$results`

<br>

**Decision Rule:**

-   If p $\leq \alpha$ → Reject H<sub>0</sub>\
-   If p \> $\alpha$ → Fail to Reject H<sub>0</sub>

<br>
:::

::: fragment
**Decision:**\
p = 0.430 \> 0.05 → **Fail to Reject H<sub>0</sub>**
:::
:::::
::::::::::

## Bayes Factor vs. Effect Size `csf()` \| `csi()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
**`r h2$results`**
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

::::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 205px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

:::: {style="text-align:center; font-size:.96rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; align-items:center; justify-content:center; line-height:1.15;"}
::: {style="font-size:.82rem; font-weight:600; margin-top:.12rem;"}
Cohen's $w$ (Association) Cramer's $V$ (Departure Size)
:::
::::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
0.33–1
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
1–3
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
3–10
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .50$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Large pattern, but evidence favors** $H_0$\
Possible noisy result or very small sample
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Large-looking pattern, weak evidence**\
Promising, but still uncertain
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Potentially important association / mismatch**\
Need more evidence before being confident
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and practically meaningful**\
A strong pattern in the counts
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Large and well supported**\
Strong evidence for a substantial pattern
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .30$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate pattern, but evidence favors** $H_0$\
Could be unstable
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Moderate-looking pattern, weak evidence**\
Unclear whether it will hold up
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Some indication of a meaningful pattern**\
Not decisive yet
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real and noticeable**\
Worth taking seriously
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Meaningful and well supported**\
Good evidence and practical relevance
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem; margin-top:.2rem;"}
$w, V = .10$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Little pattern, evidence favors** $H_0$\
Probably not much happening
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Small pattern, weak evidence overall**\
Not much to conclude
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22; background:#fcfcfc;"}
**Maybe real, but still small**\
Even if true, impact is modest
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; line-height:1.22; background:#f6fbf6;"}
**Likely real, but modest**\
Useful distinction: real does not mean large
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; line-height:1.22; background:#eef8ee;"}
**Very strong evidence for a small pattern**\
Consistent, but not large
:::
:::::::::::::::::::::::::::::::

## Step 5: Does it matter?

::::: columns
::: {.column width="50%"}
<br>

**Statistical:**

-   No significant association between Finger_Preference and Cool_Time

**Practical:**

-   Effect size: $V$ \<.001 → *small effect*\
-   Bayes Factor: bf<sub>10</sub> = 0.183 → *strong evidence for H<sub>0</sub>*

**Takeaway:**

-   Little to pattern in data, evidence favors H<sub>0</sub>
-   Probably not much of an association
-   Possible sample is too small to see meaningful pattern
:::

::: {.column width="50%"}
<br>

`h2$observed`

```{r csi_obs echo:TRUE, include:TRUE}
h2$observed
```

`h1$expected`

```{r csi_exp echo:TRUE, include:TRUE}
h2$expected
```

`h1$standardized_residuals`

```{r csi_sr echo:TRUE, include:TRUE}
h2$standardized_residuals
```
:::
:::::

## What R output belongs in a presentation?

<br>

::: fragment
#### What is the practical interpretation of this analysis for a ***business audience***? Are my estimates precise enough to tell the story I outlined in part (2)?
:::

<br>

::: fragment
-   Takeaways
-   Practical Interpretation (light on jargon & stats)
-   Limited Statistical Interpretation (don't rely on p value to tell the story)
-   No mention of $H_0$/$H_A$\
:::

::: {style="position:absolute; bottom:40px; right:60px; font-size:.7em; color:#6b7280; text-align:right;"}
`r h1$results`
:::
