---
title: "Interpreting ANOVA & Simple Linear Regression"
format:
  revealjs:
    theme:
      - default
      - ../shared/styles/boland-reveal.scss
    slide-number: true
    hash: true
    center: false
    transition: fade
    controls: true
    pdf-export: true
execute:
  echo: false
---

## Steps in Hypothesis Testing

1.  [*What statistical story am I attempting to tell?*]{.muted}

::: {.callout-tip title="What are we trying to find out?"}
:::

2.  [*What have I estimated or plotted... If no uncertainty, what should the business do?*]{.muted}

::: {.callout-tip title="What do we expect to happen?"}
:::

3.  [*Which test is appropriate and what is the null hypothesis?*]{.muted}

::: {.callout-tip title="How are we going to test it?"}
:::

4.  [*What is the statistical interpretation of the test result?*]{.muted}

::: {.callout-tip title="Do we believe it?"}
:::

5.  [*What is the practical interpretation of this analysis for a business audience?*]{.muted}

::: {.callout-tip title="Does it matter?"}
:::

# ANOVA

Comparing Means Between More Than Two Categories

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(eco230r)
library(here)
library(readr)

load(system.file("data", "t_tall.rda", package = "eco230r"))
load(system.file("data", "t_wide.rda", package = "eco230r"))
load(system.file("data", "tea_tab.rda", package = "eco230r"))
load(system.file("data", "napData.rda", package = "eco230r"))

```

## Analysis of Variance (ANOVA) `ano()`

```{r run_ano echo:FALSE, include:FALSE, message=FALSE, warning=FALSE, results="hide"}
df <- napData

h8 <- df %>%
  ano(naptime ~ light)
```

:::::::::: columns
::::: {.column width="40%"}
<br>

::: fragment
### Steps 1-3: Setup

<br>

**Test:** ANOVA\
**Function:** `ano()`

<br>

**H<sub>0</sub>:** $\mu$<sub>Low</sub> = $\mu$<sub>Warm</sub> = $\mu$<sub>Harsh</sub>\
**H<sub>A</sub>:** $\mu$<sub>Low</sub> $\neq$ $\mu$<sub>Warm</sub> $\neq$ $\mu$<sub>Harsh</sub>\
**α =** 0.05
:::

<br>

::: fragment
### Run the Test

```{r show_ano, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
h8 <- df %>%
  ano(naptime ~ light)
```
:::
:::::

::: {.column width="5%"}
:::

::::: {.column width="55%"}
<br>

::: fragment
### Step 4: Do we believe it?

`h8$results`

`r h8$results`

<br>

**Decision Rule:**

-   If p $\leq \alpha$ → Reject H<sub>0</sub>\
-   If p \> $\alpha$ → Fail to Reject H<sub>0</sub>

<br>
:::

::: fragment
**Decision:**\
p = 0.025 \< 0.05 → **Reject H<sub>0</sub>**
:::
:::::
::::::::::

## Bayes Factor vs. Effect Size `ano()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
**`r h8$results`**
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

::::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 205px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

:::: {style="text-align:center; font-size:.96rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; flex-direction:column; align-items:center; justify-content:center; line-height:1.1;"}
Variance Explained

::: {style="font-size:.82rem; font-weight:600;"}
$\omega^2$, $\xi$
:::
::::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
0.33–1
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
1–3
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
3–10
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem;"}
$\omega^2, \xi = .14$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Large variance explained, but evidence favors** $H_0$\
Group differences appear large, but this may be due to noise or small sample size
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Large-looking differences, weak evidence**\
Groups look different, but we are not yet confident this pattern is real
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Potentially important group differences**\
Group membership may explain a meaningful portion of the outcome, but evidence is still limited
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real and meaningful differences**\
Group membership explains a substantial portion of variation in the outcome
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Large and well supported**\
Strong evidence that group differences explain a large share of the outcome
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem;"}
$\omega^2, \xi = .06$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Moderate variance explained, but evidence favors** $H_0$\
There may be some group differences, but the data do not support them clearly
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Moderate differences, weak evidence**\
Groups show some separation, but it may not hold up with more data
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Some indication of meaningful differences**\
Group membership may explain part of the outcome, but not decisively
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real and noticeable differences**\
Group differences explain a meaningful portion of variation
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Meaningful and well supported**\
Good evidence that group differences matter for the outcome
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem;"}
$\omega^2, \xi = .01$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Little variance explained, evidence favors** $H_0$\
Group differences do not explain much of the outcome
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Small differences, weak evidence**\
Groups look similar, and there is little evidence of meaningful separation
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Maybe real, but small impact**\
Even if group differences exist, they explain very little of the outcome
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real, but modest differences**\
Group differences are consistent, but explain only a small portion of variation
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Very strong evidence for small differences**\
Group differences are real, but they explain very little of the outcome
:::
:::::::::::::::::::::::::::::::

## Step 5: Does it matter?

::::: columns
::: {.column width="60%"}
<br>

**Statistical:**

-   Statistically significant difference between groups

**Practical:**

-   Effect size: $\omega^2$ = 0.354 → *large effect*\
-   Bayes Factor: bf<sub>10</sub> = 3.071 → *Some evidence for H<sub>A</sub>*

**Takeaway:**

-   Likely real and meaningful differences between groups
-   Group membership explains a substantial portion of variation
-   **Post-hoc analysis must be done for individual group differences**
-   Difference between Low Light and Harsh light only significant difference (p<.05)
:::

::: {.column width="40%"}
<br>

`h8$descriptive_statistics`

```{r ano_desc echo:TRUE, include:TRUE}
h8$descriptive_statistics
```

`h8$post_hoc_analysis`

```{r ano_pha echo:TRUE, include:TRUE}
h8$post_hoc_analysis
```

:::
:::::

# Simple Linear Regression

Testing if there is a linear relationship between two ratio scales

## Simple Linear Regression `slr()`

```{r run_slr echo:FALSE, include:FALSE, message=FALSE, warning=FALSE, results="hide"}
df <- napData

h9 <- df %>%
  slr(naptime ~ timestamp)
```

:::::::::: columns
::::: {.column width="40%"}
<br>

::: fragment
### Steps 1-3: Setup

<br>

**Test:** Simple Linear Regression\
**Function:** `slr()`

<br>

**H<sub>0</sub>:** $\beta_1$ = 0\
**H<sub>A</sub>:** $\beta_1$ $\neq$ 0\
**α =** 0.05
:::

<br>

::: fragment
### Run the Test

```{r show_slr, echo=TRUE, results="hide", message=FALSE, warning=FALSE}
h9 <- df %>%
  slr(naptime ~ timestamp)
```
:::
:::::

::: {.column width="5%"}
:::

::::: {.column width="55%"}
<br>

::: fragment
### Step 4: Do we believe it?

`h9$results`

`r h9$results`

<br>

**Decision Rule:**

-   If p $\leq \alpha$ → Reject H<sub>0</sub>\
-   If p \> $\alpha$ → Fail to Reject H<sub>0</sub>

<br>
:::

::: fragment
**Decision:**\
p = 0.024 \< 0.05 → **Reject H<sub>0</sub>**
:::
:::::
::::::::::

## Bayes Factor vs. Effect Size `slr()`

:::::: {style="display:grid; grid-template-columns: .9fr 1.1fr; column-gap:1.2rem; align-items:center; margin-bottom:.25rem;"}
<!-- LEFT: test output -->

:::: {style="display:flex; justify-content:flex-start; align-items:center;"}
::: {style="text-align:left; font-size:1rem; color:black; font-weight:400; line-height:1.35; margin-left:.2rem;"}
**`r h9$results`**
:::
::::

<!-- RIGHT: Bayes factor anchor -->

::: {style="display:flex; justify-content:flex-end; align-items:center;"}
![](../shared/images/bayes_factor.svg){fig-alt="Bayes factor scale showing evidence from strong evidence for H0 through anecdotal evidence to strong and decisive evidence for HA." style="width:100%; max-width:1250px; display:block;"}
:::
::::::

::::::::::::::::::::::::::::::: {style="display:grid; grid-template-columns: 205px repeat(5, 1fr); gap:.28rem; align-items:stretch;"}
<!-- header row -->

:::: {style="text-align:center; font-size:.96rem; font-weight:700; padding:.45rem .3rem; background:#eef2f7; display:flex; flex-direction:column; align-items:center; justify-content:center; line-height:1.1;"}
Variance Explained

::: {style="font-size:.82rem; font-weight:600;"}
$R^2$
:::
::::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
$BF_{10} < 0.33$
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
0.33–1
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#dfe7f4;"}
1–3
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#d6e2f6;"}
3–10
:::

::: {style="text-align:center; font-size:.9rem; font-weight:700; padding:.45rem .3rem; background:#ccd9f2;"}
$> 10$
:::

<!-- Large -->

:::: {style="background:#2e8b57; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Large

::: {style="font-size:.95rem;"}
$R^2 = .26$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Large variance explained, but evidence favors** $H_0$\
The predictor appears important, but this may be due to noise or limited data
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Strong-looking relationship, weak evidence**\
The predictor explains a lot of variation, but we are not yet confident it is real
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem; line-height:1.22;"}
**Potentially important predictor**\
The predictor may explain a large portion of the outcome, but evidence is still limited
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real and meaningful relationship**\
The predictor explains a substantial share of variation in the outcome
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Large and well supported**\
Strong evidence that the predictor is highly informative for the outcome
:::

<!-- Medium -->

:::: {style="background:#7fc97f; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Medium

::: {style="font-size:.95rem;"}
$R^2 = .13$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Moderate variance explained, but evidence favors** $H_0$\
The predictor may matter, but the data do not support it clearly
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Moderate relationship, weak evidence**\
The predictor explains some variation, but this may not hold up
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Some indication of a useful predictor**\
The predictor may explain part of the outcome, but not decisively
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real and noticeable relationship**\
The predictor explains a meaningful portion of variation
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Meaningful and well supported**\
Good evidence that the predictor contributes to the outcome
:::

<!-- Small -->

:::: {style="background:#cfe8cf; color:white; font-weight:700; text-align:center; padding:.7rem .45rem; display:flex; flex-direction:column; justify-content:center; min-height:92px;"}
Small

::: {style="font-size:.95rem;"}
$R^2 = .02$
:::
::::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Little variance explained, evidence favors** $H_0$\
The predictor does not explain much of the outcome
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Small relationship, weak evidence**\
The predictor explains very little variation
:::

::: {style="border:1px solid #d0d0d0; padding:.55rem; font-size:.78rem;"}
**Maybe real, but small impact**\
Even if real, the predictor has limited influence on the outcome
:::

::: {style="border:1px solid #c7d9c7; padding:.55rem; font-size:.78rem; background:#f6fbf6;"}
**Likely real, but modest relationship**\
The predictor consistently explains a small portion of variation
:::

::: {style="border:1px solid #9fc49f; padding:.55rem; font-size:.78rem; background:#eef8ee;"}
**Very strong evidence for a small relationship**\
The predictor matters, but only explains a small share of the outcome
:::
:::::::::::::::::::::::::::::::

## Step 5: Does it matter?

::::: columns
::: {.column width="60%"}
**Statistical:**

-   There is a statistically significant linear relationship between timestamp and naptime.

<br>

**Practical:**

-   Effect size: $R^2$ = 0.334 → *large effect*\
-   Bayes Factor: bf<sub>10</sub> = 2.861 → *anecdotal evidence for H<sub>A</sub>*

<br>

**Takeaway:**

-   Potentially important predictor 
-   Evidence is still limited
-   Relationship described by h9$linear_regression_model
:::

::: {.column width="40%"}
<br>

`h9$linear_regression_model`

```{r slr_mod echo:TRUE, include:TRUE}
h9$linear_regression_model
```

`h9$coefficients`

```{r slr_coef echo:TRUE, include:TRUE}
h9$coefficients
```

:::
:::::



## What R output belongs in a presentation?

<br>

::: fragment
#### What is the practical interpretation of this analysis for a ***business audience***? Are my estimates precise enough to tell the story I outlined in part (2)?
:::

<br>

::: fragment
-   Takeaways
-   Practical Interpretation (light on jargon & stats)
-   Limited Statistical Interpretation (don't rely on p value to tell the story)
-   No mention of $H_0$/$H_A$\
:::

::: {style="position:absolute; bottom:40px; right:60px; font-size:.7em; color:#6b7280; text-align:right;"}
`r h8$results`
:::
