---
title: "Measures of Variability"
subtitle: "Range, interquartile range, and standard deviation"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: false
execute:
  echo: true
  warning: false
  message: false
---

## Why spread matters

Two groups can have the same average and still behave very differently. Measures of variability describe how tightly observations cluster and how far they extend.

This example uses minutes of main sleep from the ECO 230 sleep-tracking data.

```{r}
#| label: load-main-sleep
#| echo: false
sleep <- read.csv(
  here::here("week01", "data", "sleep_tracking_Jan-Jun_2025_with_zip.csv"),
  check.names = FALSE
)
x <- sleep$TotalMinutes[sleep$SleepEventType == "Main" & !is.na(sleep$TotalMinutes)]

summary_values <- data.frame(
  Measure = c("Number of records", "Minimum", "First quartile", "Median",
              "Third quartile", "Maximum", "Range", "IQR",
              "Mean", "Sample standard deviation"),
  Value = c(length(x), min(x), unname(quantile(x, .25)), median(x),
            unname(quantile(x, .75)), max(x), diff(range(x)), IQR(x),
            mean(x), sd(x))
)
knitr::kable(summary_values, digits = 1)
```

## Range

The range is the distance from the smallest observation to the largest:

\[
\text{Range} = \text{Maximum} - \text{Minimum}
\]

```{r}
diff(range(x))
```

The range is easy to explain, but it uses only two observations. A single unusual value can change it dramatically.

## Interquartile range

The interquartile range (IQR) describes the width of the middle 50% of observations:

\[
\text{IQR} = Q_3 - Q_1
\]

```{r}
quantile(x, c(.25, .50, .75))
IQR(x)
```

Because it focuses on the middle half, the IQR is more resistant to unusually high or low values than the range. It is a natural companion to the median and the box plot.

## Standard deviation

Standard deviation describes typical distance from the mean. Larger values indicate more spread around the mean.

```{r}
mean(x)
sd(x)
```

R's `sd()` calculates the **sample standard deviation**, using \(n-1\) in the denominator. If the observations are the entire population of interest rather than a sample, the population calculation is:

```{r}
population_sd <- sqrt(mean((x - mean(x))^2))
population_sd
```

Always state which population you want to describe. A complete course data file is not automatically the complete real-world population.

## What an outlier changes

The next comparison adds one implausible 1,200-minute sleep record to illustrate sensitivity. This is a diagnostic example, not a recommendation to delete high values.

```{r}
x_with_outlier <- c(x, 1200)

outlier_comparison <- data.frame(
  Measure = c("Mean", "Median", "Range", "IQR", "Sample SD"),
  Original = c(mean(x), median(x), diff(range(x)), IQR(x), sd(x)),
  WithOutlier = c(mean(x_with_outlier), median(x_with_outlier),
                  diff(range(x_with_outlier)), IQR(x_with_outlier),
                  sd(x_with_outlier))
)
knitr::kable(outlier_comparison, digits = 1)
```

An unusual value may be a data-entry error, a correctly measured rare event, or evidence that the population is more varied than expected. Investigate before changing it.

## Choosing a measure

| Situation | Useful center | Useful spread | Why |
|---|---|---|---|
| Roughly symmetric, no influential outliers | Mean | Standard deviation | Both use every observation |
| Skewed or outlier-prone | Median | IQR | Both resist extreme values |
| Need to communicate full observed span | Any | Range | Shows minimum-to-maximum extent |
| Comparing distributions | Mean and median | SD, IQR, and a graph | No single number tells the whole story |

::: {.callout-warning}
## The empirical rule has conditions

The familiar 68%, 95%, and 99.7% rule applies only to distributions that are approximately normal. Do not use it simply because a mean and standard deviation are available.
:::

## Report spread with context

A useful written result includes:

- the variable and unit;
- the population or sample being described;
- the center and spread;
- the number of nonmissing observations; and
- any meaningful skew, gaps, or unusual observations visible in a graph.

## Source note

This is a new ECO 230 replacement for the University of Leicester Student Learning Development handout *Measures of variability: the range, inter-quartile range and standard deviation* (2009). The explanation and examples have been rewritten for R and current course data; outdated spreadsheet instructions were not reproduced.
