---
title: "Working with Percentages"
subtitle: "Choose the denominator before interpreting the result"
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: false
execute:
  echo: true
  warning: false
  message: false
---

## Why percentages help

Counts answer "how many?" Percentages answer "how large relative to what?" They are especially useful when groups have different totals.

Every percentage has a numerator and a denominator:

\[
\text{Percentage} = \frac{\text{part}}{\text{whole}} \times 100
\]

Name the denominator in words before calculating.

## Part of a whole

The Airbnb example below calculates each room type's share of listings. The denominator is **all listings in this Chicago data file with a recorded room type**.

```{r}
#| label: room-type-share
#| echo: false
airbnb <- read.csv(
  here::here("shared", "data", "airbnb_chicago.csv"),
  check.names = FALSE
)

room_counts <- sort(table(airbnb$room_type, useNA = "no"), decreasing = TRUE)
room_share <- data.frame(
  RoomType = names(room_counts),
  Listings = as.integer(room_counts),
  Percent = 100 * as.numeric(prop.table(room_counts))
)
knitr::kable(room_share, digits = 1)
```

If the question instead concerned only reviewed listings or only listings below a price threshold, both the numerator and denominator would change.

## Row, column, and table percentages

A crosstab can support three different questions. The example compares room type with instant-booking status.

```{r}
#| label: crosstab-percentages
booking_table <- table(
  RoomType = airbnb$room_type,
  InstantBookable = airbnb$instant_bookable,
  useNA = "no"
)
booking_table
```

- **Row percentage:** Within each room type, what percentage is instantly bookable? Use `prop.table(booking_table, 1)`.
- **Column percentage:** Within each instant-booking category, what percentage is each room type? Use `prop.table(booking_table, 2)`.
- **Table percentage:** What percentage of all listings is in each combination? Use `prop.table(booking_table)`.

```{r}
#| label: row-percentages
round(100 * prop.table(booking_table, 1), 1)
```

Do not label a table merely "percent." Say whether it contains row, column, or total percentages.

## Percentage change

Percentage change compares a change with its starting value:

\[
\text{Percentage change} =
\frac{\text{New value} - \text{Old value}}{\text{Old value}} \times 100
\]

```{r}
#| label: sleep-percentage-change
#| echo: false
sleep <- read.csv(
  here::here("week01", "data", "sleep_tracking_Jan-Jun_2025_with_zip.csv"),
  check.names = FALSE
)
main_sleep <- sleep[sleep$SleepEventType == "Main", ]
monthly_mean <- aggregate(TotalMinutes ~ Month, main_sleep, mean, na.rm = TRUE)

old_value <- monthly_mean$TotalMinutes[monthly_mean$Month == "January"]
new_value <- monthly_mean$TotalMinutes[monthly_mean$Month == "June"]
change <- 100 * (new_value - old_value) / old_value

knitr::kable(
  data.frame(
    Comparison = "Average main-sleep minutes: January to June",
    January = old_value,
    June = new_value,
    PercentChange = change
  ),
  digits = 1
)
```

A negative result indicates a decrease. The starting value is the denominator, so reversing the comparison produces a different percentage.

## Percentage points are not percent change

Suppose a response rate rises from 40% to 50%.

- It increased by **10 percentage points**: \(50\% - 40\%\).
- It increased by **25% relative to the old rate**: \((50 - 40) / 40 \times 100\).

These statements are both correct, but they answer different questions.

## Common denominator mistakes

1. Comparing counts when group sizes differ.
2. Dividing by the new value instead of the starting value for percentage change.
3. Ignoring missing values that shrink the denominator.
4. Mixing row and column percentages in one interpretation.
5. Reporting a large percentage change from a very small starting value without giving the counts.

::: {.callout-tip}
## Report counts and percentages together

Write "18 of 60 listings (30%)" rather than only "30%." The count reveals the scale and makes the denominator auditable.
:::

## Source note

This is a new ECO 230 replacement for the University of Leicester Student Learning Development handout *Working with percentages* (2009). The explanation and examples have been rewritten using current course data.
