---
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: false

execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| include: false
library(tidyverse)
library(tidyr)
library(eco230r)
library(ggplot2)
library(knitr)
library(Hmisc)
load("grading_1.RData")
load("grading_2.RData")
# I'm going to turn off scientific notation:
options(scipen = 999)
# If you want to turn it back on you can uncomment and run this line:
# options(scipen = 0)
```

## ECO 230 PRACTICUM RUBRIC AND GUIDELINES

::: {.callout-note title="Timeline"}
-   On Week 13 the prompt and data description will be posted on Canvas in your Final Exam Preparation Module
-   The data will be released during our final exam period.
-   You will have 120 minutes to complete both the traditional portion of the final and the Practicum.
:::

::: {.callout-note title="Requirements"}
Your analysis should include:

-   At least one graph
-   At least one table (e.g. pivot table)
-   At least one hypothesis test
-   A brief bullet point summary of your findings
-   Work in Posit.cloud, Excel, and Tableau (optional) - **Posit.cloud and Excel are required Tableau is optional.**

***You do not need to address all questions in the prompt. You may choose to focus on only one if you like.***
:::

::: {.callout-note title="Submitting Completed Work"}
Upload your knit RMarkdown pdf, your Excel workbook (.xlsx) and/or your Tableau Packaged Workbook (.twbx) to Canvas.
:::

::: {.callout-note title="Permitted and Forbidden Resources"}
You are welcome to use:

-   Your own hand written notes
-   Prior work in RStudio including completed case work, homework, and practice assignments
-   Any web resources including Primers from Posit.cloud
-   Ideas developed while discussing the case with colleagues before the exam

You may not:

-   Discuss work with colleagues during the exam via any channel (no text, email, chat etc.)
-   Count on your instructor for substantive help during the exam or ask for comments on planned analysis before the exam.
-   Use text generating AI. We all love robots and their plans for world domination, but your answers and analysis should be in your own words.
:::

::: {.callout-note title="Suggestions for Successful Work"}
-   Plan ahead! Know what you intend to make before you launch the final and choose what test will be appropriate.
-   Keep it simple. Do not attempt to do more than is required of you. Do not make graphs with more than 2 variables!
-   Do not forget what you've learned about data visualization. No 3D. Use color to convey meaning only.
-   Pay attention to measures vs. dimensions and green vs. blue pills in Tableau.
-   Do your hypothesis test in R.
-   Don't forget to write your bullet point interpretations!
-   Save your work correctly: .xlsx instead of .csv, .twbx instead of .twb, .Rmd knit to .pdf.
:::

::: {.callout-note title="Formal Grading Criteria: The Practicum Counts for 40% of Your Final Exam Grade"}
```{r}
kable(grading_1)
```

```{r}
kable(grading_2)
```
:::