---
title: "Homework 5: Hypothesis Tests"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true

execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| include: false
library(dplyr)
library(eco230r)
```

## HOMEWORK 5: HYPOTHESIS TESTS

::: {.callout-note title="Objectives"}
-   Complete step 3 of the hypothesis testing process: Which test is appropriate and what is the null hypothesis?
-   Choose an appropriate inferential test for each hypothesis.
-   Translate your hypothesis to null ($H_0$) and alternate ($H_A$) form.
-   Draft pseudocode that outlines how you would run each test in R.
:::

::: {.callout-important title="Submission Instructions"}
1.  Open the Posit Cloud project space: https://posit.cloud/spaces/3173/content/5508040\
2.  Open the file named **Homework_5.Rmd**.\
3.  Complete Hypotheses 1–3 (fill in hypothesis, test type, $H_0$, $H_A$, and pseudocode).\
4.  Knit to .pdf and submit the knitted file.

**Important** - Running code is NOT required for this homework. - If you try code and it prevents knitting, comment it out using #.
:::

::: {.callout-note title="What You Will Likely Need To Do In R"}
Most hypotheses will require:

-   Load data
-   Filter to relevant subset (example: IL, MI, WI)
-   Run the appropriate test

Example pseudocode (ANOVA across IL/MI/WI):

```{r}
#| eval: false
# Load data

# Filter to states

# Run ANOVA

```
:::

------------------------------------------------------------------------

### Hypothesis 1

**What statistical story am I attempting to tell?**

Hypothesis:

**Which test is appropriate and what is the null hypothesis?**

Test Type:

$H_0$:

$H_A$:

**Pseudocode**

```{r}
#| eval: false
# Load data
# Filter (if needed)
# Run test
```

------------------------------------------------------------------------

### Hypothesis 2

**What statistical story am I attempting to tell?**

Hypothesis:

**Which test is appropriate and what is the null hypothesis?**

Test Type:

$H_0$:

$H_A$:

**Pseudocode**

```{r}
#| eval: false
# Load data
# Filter (if needed)
# Run test
```

------------------------------------------------------------------------

### Hypothesis 3

**What statistical story am I attempting to tell?**

Hypothesis:

**Which test is appropriate and what is the null hypothesis?**

Test Type:

$H_0$:

$H_A$:

**Pseudocode**

```{r}
#| eval: false
# Load data
# Filter (if needed)
# Run test
```
