---
title: "Homework 5: Hypothesis-Test Planning"
subtitle: "In-Person Sections"
format:
  typst:
    toc: false
    papersize: us-letter
    fontsize: 10pt
    linestretch: 1
    margin:
      x: 1in
      y: 1in
    number-sections: false
    page-numbering: "1"
execute:
  echo: true
  warning: false
  message: false
editor: source
---

::: {.callout-note appearance="minimal" icon="false" title="Purpose"}
This assignment focuses on Step 3 of the five-step hypothesis-testing process:
selecting a test and translating a statistical story into null and alternative
hypotheses.

You will choose three project hypotheses, select an appropriate inferential test
for each one, and write pseudocode that outlines how the test would run in R.
:::

## Objectives

By the end of this assignment, you will be able to:

- distinguish among the inferential tests introduced in the course;
- match a test to the variables and study design;
- state $H_0$ and $H_A$ in context; and
- outline data-loading, filtering, and testing steps as R pseudocode.

## Resources

- Week 6 slides and test-selection resources
- Your project analysis plan and three candidate hypotheses
- The four `.csv` files in the `data` folder
- `eco230r` function examples from lecture

::: {.callout-tip appearance="minimal" icon="false" title="Expectations"}
Running code is not required. Your pseudocode must show a reasonable path from
the dataset to the selected test. If you experiment with code that prevents the
document from rendering, comment it out with `#`.
:::

## 1) Review the Available Tests

| Function | Test | Typical structure |
|---|---|---|
| `csf()` | Chi-square goodness-of-fit test | One categorical variable |
| `csi()` | Chi-square test of independence | Two categorical variables |
| `ost()` | One-sample t-test | One numeric variable vs a reference value |
| `idt()` | Independent-samples t-test | Numeric outcome by two independent groups |
| `idw()` | Wilcoxon rank-sum test | Numeric/ordinal outcome by two independent groups |
| `pst()` | Paired-samples t-test | Numeric outcome measured in pairs |
| `psw()` | Wilcoxon signed-rank test | Numeric/ordinal outcome measured in pairs |
| `ano()` | One-way ANOVA | Numeric outcome by three or more groups |
| `slr()` | Simple linear regression | Numeric outcome and one numeric predictor |

Select a test based on the variables, number of groups, independence or pairing,
and the type of comparison in your statistical story.

---

## 2) Review the Worked Example

### Statistical story

The average number of bedrooms in Chicago Airbnb listings differs from two.

### Estimate or plot

A descriptive summary of `bedrooms` would show whether the sample mean is above
or below two. This estimate provides direction but not evidence about sampling
uncertainty.

### Test and hypotheses

The outcome is one numeric variable compared with a reference value, so the
appropriate test is a one-sample t-test using `ost()`.

$H_0$: The population mean number of bedrooms is 2.

$H_A$: The population mean number of bedrooms is not 2.

$\alpha = 0.05$

```{r}
#| eval: false
df <- read.csv(file.path("data", "airbnb_chicago.csv"))

h1 <- df |>
  ost(~ bedrooms, mu = 2)

print(h1)
```

The exact argument for the reference value may depend on the current `eco230r`
function interface. At this stage, the important work is identifying the test,
hypotheses, and intended code sequence.

Steps 4-5, the statistical and practical interpretations, are not required for
this homework because you are planning rather than interpreting test results.

---

## 3) Plan Hypothesis 1

1. **Statistical story:** What relationship, difference, or pattern are you
   investigating, and why might it matter?

   *Your response:*



2. **Estimate or plot:** What descriptive estimate or plot would support this
   story? If there were no uncertainty, what action would the business take?

   *Your response:*



3. **Test and hypotheses:** Which test is appropriate? State $H_0$, $H_A$,
   and the alpha level.

   **Test and function:**

   $H_0$:

   $H_A$:

   $\alpha = 0.05$

```{r}
#| eval: false
# Load the selected dataset.

# Filter or derive variables if needed.

# Run the selected eco230r test.
```

---

## 4) Plan Hypothesis 2

1. **Statistical story:** What relationship, difference, or pattern are you
   investigating, and why might it matter?

   *Your response:*



2. **Estimate or plot:** What descriptive estimate or plot would support this
   story? If there were no uncertainty, what action would the business take?

   *Your response:*



3. **Test and hypotheses:** Which test is appropriate? State $H_0$, $H_A$,
   and the alpha level.

   **Test and function:**

   $H_0$:

   $H_A$:

   $\alpha = 0.05$

```{r}
#| eval: false
# Load the selected dataset.

# Filter or derive variables if needed.

# Run the selected eco230r test.
```

---

## 5) Plan Hypothesis 3

1. **Statistical story:** What relationship, difference, or pattern are you
   investigating, and why might it matter?

   *Your response:*



2. **Estimate or plot:** What descriptive estimate or plot would support this
   story? If there were no uncertainty, what action would the business take?

   *Your response:*



3. **Test and hypotheses:** Which test is appropriate? State $H_0$, $H_A$,
   and the alpha level.

   **Test and function:**

   $H_0$:

   $H_A$:

   $\alpha = 0.05$

```{r}
#| eval: false
# Load the selected dataset.

# Filter or derive variables if needed.

# Run the selected eco230r test.
```

## Submission

Render this document and submit `Homework_05_IP.pdf` in Canvas. The PDF must
contain three statistical stories, three test selections, contextual null and
alternative hypotheses, and pseudocode for each test.

## Checklist

- [ ] I selected three hypotheses connected to my project.
- [ ] Each test matches the variables, groups, and study design.
- [ ] Each $H_0$ and $H_A$ is stated in context.
- [ ] Each pseudocode block shows loading, preparation, and testing steps.
- [ ] Any incomplete experimental code is commented out.
- [ ] I reviewed and submitted `Homework_05_IP.pdf`.

**Grading focus:** Your work will be evaluated for appropriate test selection,
testable hypotheses, clear context, and pseudocode that reflects a feasible
analysis plan.
